---
title: "Sample"
author: "Ginnie Ma"
date: "6/29/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulating fake data

Using mvrnorm, simulate normal data with mean of 0 and given covariance matrix, Sigma. Checked by calculating covariance matrix of simulated data to match up with inputed covariance. Used below as error matrix when generating fake data.

```{r mvrnorm_1000samples}
set.seed(1234)
mu=rep(0, 3)
Sigma=2*matrix(c(1, 0.7, 0.3, 0.7, 1, 0.7, 0.3, 0.7, 1), 3, 3)
data=mvrnorm(10000, mu, Sigma)
print("Correlation")
cor = cor(data)
cor
#print("Variance")
#var(data)
print("Covariance")
cov=cov(data)
cov
```

Random "filler" W and H
```{r W and H data}
{
  #W matrix: signatures of pure tumor cells
  #H matrix: how much tumor in each sample
  set.seed(1234)
  #3 x 4 matrix
  k = 4
  W = matrix(c(1+rnorm(k, 0, 1), 2+rnorm(k, 0, 1), 3+rnorm(k, 0, 1)), nrow=3, byrow=TRUE)
  H = matrix(runif(k*1000, 0, 1), nrow=k, byrow=TRUE)
}
```

Parameters:
G = number of genes
S = number of samples
W - g by k matrix, library of signatures of pure tumor cells
H - k by s matrix, how much tumor is in each sample
Sigma - g by g matrix, covariance matrix for random normal data generated
mu - vector of g elements, means of random normal data generated

Create fake data, G x S matrix, composed of WH + error (random normal data)

```{r Simulate data}
mu = rep(0, 3)
Sigma=2*matrix(c(1, 0.7, 0.3, 0.7, 1, 0.7, 0.3, 0.7, 1), 3, 3)
simulateData <- function(W, H, Sigma, mu)
{
  G = nrow(W)
  S = ncol(H)
  set.seed(1234)
  # simulate the random normal data which serves as error
  error = mvrnorm(S, mu, Sigma)
  data = W %*% H + t(error)
  return (data)
}
```
PCA plot


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
